---
title: "Exp 001-CO261 karyotapR analysis"
author: "Joey Mays"
date: '2023-01-14'
output:
  pdf_document:
    toc: yes
  html_notebook:
    code_folding: show
    toc: yes
    toc_float:
      collapsed: no
      smooth_scroll: yes
    theme: flatly
    df_print: paged
  html_document:
    code_folding: show
    toc: yes
    toc_float:
      collapsed: no
      smooth_scroll: yes
    theme: flatly
    df_print: paged
editor_options:
  chunk_output_type: inline
  markdown:
    wrap: 72
---

# Setup

```{r message=FALSE}
library(karyotapR)
library(tidyverse)
library(ggplot2)
library(ggdist)
library(here)
library(ComplexHeatmap)
```

```{r}
set.seed(20221210)
```

```{r}
here::i_am("analysis-notebooks/exp01-karyotapr-analysis.Rmd")
```

# Load Data

```{r}
tapestri.filename <- here("datasets", "tapestri-experiment01-panelv1.h5")
```

# Construct Object

```{r}
exp1 <- createTapestriExperiment(tapestri.filename, panel.id = "CO261")
```

# AF Clustering

## Embedding

```{r}
exp1 <- runPCA(exp1, "alleleFrequency", sd.min.threshold = 20)
PCAKneePlot(exp1)
```


```{r}
reducedDimPlot(exp1, dim.reduction = "pca")
exp1 <- runUMAP(exp1, alt.exp = "alleleFrequency", use.pca.dims = T, pca.dims = 1:4)
```

```{r}
reducedDimPlot(exp1, alt.exp = "alleleFrequency", dim.reduction = "umap")
```

## Clustering

```{r}
exp1 <- runClustering(exp1, alt.exp = "alleleFrequency", eps = 0.3)
```


```{r}
reducedDimPlot(exp1, dim.reduction = "umap", group.label = "cluster")
```

Number of cells in each cluster

```{r}
fct_count(colData(exp1)$cluster)
```
Filter

```{r}
exp1.subset <- exp1[,colData(exp1)$cluster %in% 1:5]
```

```{r}
reducedDimPlot(exp1.subset, dim.reduction = "umap", group.label = "cluster")
```


## Probes

```{r}
summary(rowData(exp1.subset)$median.reads > 0)
rowData(exp1.subset)[rowData(exp1.subset)$median.reads == 0,]
```
```{r}
exp1.subset <- exp1.subset[rowData(exp1.subset)$median.reads > 0,] #8 filtered for exp 1
```

# Normalization

```{r}
exp1.subset <- calcNormCounts(exp1.subset)
```

## Copy Number Calculation

Rename cluster labels

```{r}
colData(exp1.subset)$cluster <- fct_recode(colData(exp1.subset)$cluster, cellline1 = "1", cellline2 = "2", cellline3 = "3", cellline4 = "4", cellline5 = "5")
colData(exp1.subset)$cluster <- fct_drop(colData(exp1.subset)$cluster)
fct_count(colData(exp1.subset)$cluster)
```

Following probes have control cell median equal to 0. Filter out prior to proceeding.

```{r}
amps.to.exlcude <- c("AMPL158845", "AMPL147144", "AMPL147154", "AMPL135088", "AMPL159975",
                     "AMPL147293", "AMPL113086", "AMPL147323", "AMPL159221", "AMPL159339", 
                     "AMPL158390", "AMPL158655", "AMPL147043")
exp1.subset <- exp1.subset[!rownames(exp1.subset) %in% amps.to.exlcude,]
```

Generate control.copy.number lookup table for RPE1 cells, set 10q to ploidy = 3

```{r}
control.copy.number <- generateControlCopyNumberTemplate(exp1.subset, sample.feature.label = "cellline2", copy.number = 2)
control.copy.number["chr10q", "copy.number"] <- 3
exp1.subset <- calcCopyNumber(exp1.subset, control.copy.number = control.copy.number, sample.feature = "cluster")
```

```{r}
ploidy.tidy <- getTidyData(exp1.subset, assay = "copyNumber")
```

```{r}
ploidy.tidy %>% group_by(feature.id, cluster) %>% summarize(median.ploidy = median(copyNumber), .groups = "drop") %>% 
    ggplot(aes(x=feature.id, y = median.ploidy)) + geom_point(size = 1) + facet_wrap(cluster~., ncol = 1) + theme(axis.text.x = element_blank())
```

## Copy Number Smoothing

```{r}
exp1.subset <- calcSmoothCopyNumber(exp1.subset)
rpe1.bc <- colData(exp1.subset) %>% as_tibble() %>% filter(cluster == "cellline2") %>% pull(cell.barcode)
exp1.subset <- calcGMMCopyNumber(exp1.subset, cell.barcodes = rpe1.bc, control.copy.number = control.copy.number, model.components = 1:6)
```

```{r}
assayHeatmap(exp1.subset, assay = "copyNumber", split.col.by = "chr", split.row.by = "cluster", color.preset = "copy.number.denoise")
```

```{r}
colData(exp1.subset)$cluster <- fct_recode(colData(exp1.subset)$cluster, LS513 = "cellline1", RPE1 = "cellline2", SW48 = "cellline3", LoVo = "cellline4", CL11 = "cellline5")
```

```{r}
smoothed.ploidy.tidy.chr <- getTidyData(exp1.subset, alt.exp = "smoothedCopyNumberByChr", assay = "smoothedCopyNumber")
```

```{r}
ggplot(smoothed.ploidy.tidy.chr, aes(x = feature.id, y = smoothedCopyNumber)) + facet_wrap(cluster ~., ncol = 1) +
    ggdist::stat_halfeye(point_size = 2, alpha = 0.8)
```

```{r}
smoothed.ploidy.tidy.arm <- getTidyData(exp1.subset, alt.exp = "smoothedCopyNumberByArm", assay = "smoothedCopyNumber")
```

```{r}
ggplot(smoothed.ploidy.tidy.arm, aes(x = feature.id, y = smoothedCopyNumber)) + facet_wrap(cluster ~., ncol = 1) +
    ggdist::stat_halfeye(point_size = 2, alpha = 0.8) + theme(axis.text.x = element_text(angle = 45, hjust = 1)) + 
    scale_y_continuous(breaks = c(0,2,4,6,8,10))
```

# Heatmaps

```{r}
cellline.anno <- rowAnnotation(cellline = colData(exp1.subset)$cluster, annotation_name_side = "top", annotation_name_gp = grid::gpar(fontsize = 6),
                               col = list(cellline = c("RPE1" = "#785EF0", "LS513" = "#648FFF", "SW48" = "#DC267F", "LoVo" = "#FE6100", "CL11" = "#FFB000")))
```

Unbiased clustering of all cell lines. 
Excluding chr 16-22 to reduce noise

```{r}
exp1.subset.matrix <- scale(t(assay(altExp(exp1.subset, "smoothedCopyNumberByArm"), "smoothedCopyNumber"))[,c(1:27, 40, 41)])
wss <- numeric(10)
cluster.list <- list()

for (i in seq_along(wss)) {
  km.result <- kmeans(exp1.subset.matrix, centers = i, nstart = 10)
  cluster.list[[i]] <- km.result$cluster
  wss[i] <- km.result$tot.withinss
}

ggplot(data.frame(clusters = seq_along(wss), wss = wss), aes(x = clusters, y = wss, group = 1)) +
    geom_point() +
    geom_line()

assayHeatmap(exp1.subset, alt.exp = "smoothedCopyNumberByArm", assay = "smoothedCopyNumber", color.preset = "copy.number", name = "smCN", 
             left_annotation = cellline.anno, use_raster = T, row_split = cluster.list[[5]], border = F)
```

```{r}
assayHeatmap(exp1.subset, alt.exp = "smoothedCopyNumberByChr", assay = "smoothedCopyNumber", split.row.by = "cluster", color.preset = "copy.number.denoise", name = "smCN", column_names_rot = 0, column_names_centered = T)
```

```{r}
assayHeatmap(exp1.subset, alt.exp = "smoothedCopyNumberByArm", assay = "smoothedCopyNumber", split.row.by = "cluster", color.preset = "copy.number.denoise", name = "smCN")
```

```{r}
assayHeatmap(exp1.subset, alt.exp = "smoothedCopyNumberByChr", assay = "gmmCopyNumber", split.row.by = "cluster", color.preset = "copy.number.denoise", name = "gmmCN", column_names_rot = 0, column_names_centered = T)
```

```{r}
assayHeatmap(exp1.subset, alt.exp = "smoothedCopyNumberByArm", assay = "gmmCopyNumber", split.row.by = "cluster", color.preset = "copy.number.denoise", name = "gmmCN")
```

## Clustering

### LoVo

LoVo appears to have subclones with exclusive events in 5p and 15q

```{r}
exp1.subset.lovo <- exp1.subset[,colData(exp1.subset)$cluster == "LoVo"]
```

```{r}
#subsetting only chr 5p, 15q 
lovo.subset2 <- t(assay(altExp(exp1.subset.lovo, "smoothedCopyNumberByArm"), "smoothedCopyNumber"))[,c("chr5p", "chr15q")]
wss <- numeric(10)
cluster.list <- list()

for (i in 1:10) {
  km.result <- kmeans(lovo.subset2, centers = i, nstart = 10)
  cluster.list[[i]] <- km.result$cluster
  wss[i] <- km.result$tot.withinss
}

ggplot(data.frame(clusters = 1:10, wss = wss), aes(x = clusters, y = wss, group = 1)) +
    geom_point() +
    geom_line()
```

```{r}
assayHeatmap(exp1.subset.lovo, alt.exp = "smoothedCopyNumberByArm", assay = "smoothedCopyNumber", split.row.by = "cluster", color.preset = "copy.number.denoise", name = "smCN", 
             row_split = cluster.list[[2]])
assayHeatmap(exp1.subset.lovo, alt.exp = "smoothedCopyNumberByArm", assay = "smoothedCopyNumber", split.row.by = "cluster", color.preset = "copy.number.denoise", name = "smCN", 
             row_split = cluster.list[[3]])

#split by chr15q value
assayHeatmap(exp1.subset.lovo, alt.exp = "smoothedCopyNumberByArm", assay = "smoothedCopyNumber", split.row.by = "cluster", color.preset = "copy.number.denoise", name = "smCN", 
             row_split = ifelse(lovo.subset2[,"chr15q"] > 2.8, 1, 2))
```

```{r}
as.data.frame(lovo.subset2) %>% add_column(k2 = as.factor(cluster.list[[2]]), k3 = as.factor(cluster.list[[3]])) %>% 
    pivot_longer(values_to = "cn", names_to = "chr", cols = starts_with("chr")) %>% 
    ggplot(aes(x = cn)) +
    geom_density() +
    facet_grid(k2 ~ chr)

as.data.frame(lovo.subset2) %>% add_column(k2 = as.factor(cluster.list[[2]]), k3 = as.factor(cluster.list[[3]])) %>% 
    pivot_longer(values_to = "cn", names_to = "chr", cols = starts_with("chr")) %>% 
    ggplot(aes(x = cn)) +
    geom_density() +
    facet_grid(k3 ~ chr)

as.data.frame(lovo.subset2) %>% add_column(cluster = ifelse(lovo.subset2[,"chr15q"] > 2.8, 1, 2)) %>% 
    pivot_longer(values_to = "cn", names_to = "chr", cols = starts_with("chr")) %>% 
    ggplot(aes(x = cn)) +
    geom_density() +
    facet_grid(cluster ~ chr)
```

# Label Transfers

# Save Objects

```{r}
saveRDS(exp1, file = here("datasets", "exp01.full.tapestriexperiment.RDS"))
saveRDS(exp1.subset, file = here("datasets", "exp01.tapestriexperiment.RDS"))
```

# session_info

```{r}
sessioninfo::session_info()
```
