---
title: "Figure D Simluation"
author: "Joey Mays"
date: '2023-01-16'
output:
  pdf_document:
    toc: yes
  html_notebook:
    code_folding: show
    toc: yes
    toc_float:
      collapsed: no
      smooth_scroll: yes
    theme: flatly
    df_print: paged
  html_document:
    code_folding: show
    toc: yes
    toc_float:
      collapsed: no
      smooth_scroll: yes
    theme: flatly
    df_print: paged
editor_options:
  chunk_output_type: inline
  markdown:
    wrap: 72
---

# Setup

```{r message=FALSE}
library(karyotapR)
library(tidyverse)
library(ggplot2)
library(here)
library(furrr)
```

```{r}
set.seed(20221210)
#plan(multicore, workers = parallel::detectCores()-1)
plan(multisession, workers = 4)
```

```{r}
here::i_am("analysis-notebooks/Figure-D-downsample-sim.Rmd")
```

```{r}
source(here("scripts", "gmm-calc-TPR.R"))
```

```{r}
dir.create(path = here("figures", "figure-d"), showWarnings = F)
```

# Load Data

```{r}
exp1 <- readRDS(file = here("datasets", "exp01.tapestriexperiment.RDS"))
```

# Analysis

## Probe Downsampling

### Chr 2

Subset TapestriExperiment to only RPE1 cells and chr2. 

```{r}
rpe1.bcs <- colData(exp1) %>% as_tibble() %>% filter(cluster == "RPE1") %>% pull(cell.barcode)
chr2.probes <- rowData(exp1) %>% as_tibble() %>% filter(chr == 2) %>% pull(probe.id)
rpe1 <- exp1[chr2.probes, rpe1.bcs]
```

Repeatedly sample probes. Start at 22 with 50 reps, add one instance of 24. 

```{r}
n.reps <- 50 # n reps per probe level
probe.levels <- seq(22, 4, -2) # sequence of probe levels
downsample.sequence <- rep(probe.levels, each = n.reps)
downsample.sequence <- c(23, downsample.sequence)
```

```{r}
rpe1.probe.samples <- list()
for(i in seq_len(length(downsample.sequence))){
        rpe1.probe.samples[[i]] <- sample(chr2.probes, size = downsample.sequence[i])
}
```

```{r}
sampled.matrix.list <- list()
rpe1.counts <- assay(rpe1, "counts")
rpe1.copynumber <- assay(rpe1, "copyNumber")
for(i in seq_len(length(rpe1.probe.samples))){
  
    current.mat <- rpe1.counts[rpe1.probe.samples[[i]], ]
    
    row.data <- data.frame(probe.id = rpe1.probe.samples[[i]], chr = 2, 
                           start.pos = 0, end.pos = 100, row.names = rpe1.probe.samples[[i]])
  
  te <- karyotapR:::.createTapestriExperiment.sim(counts = current.mat, probe.metadata = row.data)
  assay(te, "counts") <- current.mat
  assay(te, "copyNumber") <- rpe1.copynumber[rpe1.probe.samples[[i]], ]
  
   sampled.matrix.list[[i]] <- te
}
```

Simulation Function

```{r}
sim.step <- function(tapestri.object.init, step, n.reps){
  
    control.cn <- generateControlCopyNumberTemplate(tapestri.object.init, copy.number = 2, sample.feature.label = "dummy.val")
    #control.cn["chr10q", "copy.number"] <- 3 
    colData(tapestri.object.init)$dummy.feature <- "dummy.val"
    
    tapestri.object <- suppressMessages(calcSmoothCopyNumber(tapestri.object.init, 
                                                             method = "weighted.median", 
                                                             control.copy.number = control.cn, 
                                                             sample.feature = "dummy.feature")) #smooth copy number

    tapestri.object <- suppressMessages(calcGMMCopyNumber(tapestri.object, cell.barcodes = colnames(tapestri.object), control.cn, model.components = 1:6))
    
    gmm.metrics <- calcTruePosRate(tapestri.object) %>% filter(cn.sim.class != "sim_cn6") %>% 
      dplyr::select(feature.id, cn.sim.class, true.pos.p)
    
    gmm.metrics <- gmm.metrics %>% add_column(sample.size = nrow(tapestri.object))
    
        step.local <- step %% n.reps
    gmm.metrics$replicate <- ifelse(step.local == 0, n.reps, step.local) 
    
    gmm.metrics$actual.sensitivity <- getTidyData(tapestri.object, "smoothedCopyNumberByChr", "gmmCopyNumber") %>% mutate(cn.concordance = gmmCopyNumber == 2) %>% 
        count(cn.concordance) %>% mutate(pct.concordance = (n*100)/sum(n)) %>% filter(cn.concordance == TRUE) %>% pull(pct.concordance)
  
    return(gmm.metrics)
}
```

Running GMM for components 1:6 so cn=5 has an upper boundary but only evaluating 1:5.

```{r}
sim.results.chr2 <- future_imap(sampled.matrix.list, ~sim.step(tapestri.object.init = .x, step = .y, n.reps = n.reps), .progress = T, .options = furrr_options(seed = 20221210))
sim.results.chr2 <- list_rbind(sim.results.chr2)
```

```{r}
saveRDS(sim.results.chr2, file = here("datasets", "chr2.downsampleProbesSim.results.RDS"))
```

### Chr 6

Subset TapestriExperiment to only RPE1 cells and chr6.

```{r}
rpe1.bcs <- colData(exp1) %>% as_tibble() %>% filter(cluster == "RPE1") %>% pull(cell.barcode)
chr6.probes <- rowData(exp1) %>% as_tibble() %>% filter(chr == 6) %>% pull(probe.id)
rpe1 <- exp1[chr6.probes, rpe1.bcs]
```

Repeatedly sample probes. Start at 22 with 50 reps, add one instance of 24. 

```{r}
n.reps <- 50 # n reps per probe level
probe.levels <- seq(16, 4, -2) # sequence of probe levels
downsample.sequence <- rep(probe.levels, each = n.reps)
downsample.sequence <- c(18, downsample.sequence)
```

```{r}
rpe1.probe.samples <- list()
for(i in seq_len(length(downsample.sequence))){
        rpe1.probe.samples[[i]] <- sample(chr6.probes, size = downsample.sequence[i])
}
```

```{r}
sampled.matrix.list <- list()
rpe1.counts <- assay(rpe1, "counts")
rpe1.copynumber <- assay(rpe1, "copyNumber")
for(i in seq_len(length(rpe1.probe.samples))){
  
    current.mat <- rpe1.counts[rpe1.probe.samples[[i]], ]
    
    row.data <- data.frame(probe.id = rpe1.probe.samples[[i]], chr = 6, 
                           start.pos = 0, end.pos = 100, row.names = rpe1.probe.samples[[i]])
  
  te <- karyotapR:::.createTapestriExperiment.sim(counts = current.mat, probe.metadata = row.data)
  assay(te, "counts") <- current.mat
  assay(te, "copyNumber") <- rpe1.copynumber[rpe1.probe.samples[[i]], ]
  
   sampled.matrix.list[[i]] <- te
}
```

Run simulation

```{r}
sim.results.chr6 <- future_imap(sampled.matrix.list, ~sim.step(tapestri.object.init = .x, step = .y, n.reps = n.reps), .progress = T, .options = furrr_options(seed = 20221210))
sim.results.chr6 <- list_rbind(sim.results.chr6)
```

```{r}
saveRDS(sim.results.chr6, file = here("datasets", "chr6.downsampleProbesSim.results.RDS"))
```

# session_info

```{r}
sessioninfo::session_info()
```
